A fast spectral method for inelastic collision operator and the heated granular flow\tnoteref{mytitlenote}

### 2018-07-08 paper/draft/elsarticle_v2.tex
cleanup

% personal packages
\usepackage{amsmath,amssymb}
\usepackage{subfig}
\usepackage{tabu}
\usepackage{booktabs}
\usepackage{float}

@@

\title{A fast spectral method for inelastic collision operator and the heated granular flow\tnoteref{mytitlenote}}
% \tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
% \author{Jingwei Hu, Zheng Ma\fnref{myfootnote}}
% \address{Department of Mathematics, Purdue University, 150 N. University Street, West Lafayette, IN, USA}
% \fntext[myfootnote]{}

%% or include affiliations in footnotes:
\author[mymainaddress]{Jingwei Hu}
\ead{jingweihu@purdue.edu}

\author[mymainaddress]{Zheng Ma}
% \cortext[mycorrespondingauthor]{Corresponding author}
\ead{ma531@purdue.edu}

\address[mymainaddress]{150 N. University Street, West Lafayette, IN, USA}
% \address[mysecondaryaddress]{360 Park Avenue South, New York}

@@

% \begin{keyword}
% \texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
% \MSC[2010] 00-01\sep  99-00
% \end{keyword}

@@

It has been found in the past few decays that the granular gases behave fundamentally different from the usual molecular gases modelled as elastically colliding spheres. The rich phenomenology of such systems, such as the formation of clusters and shear instability, draws a lot of attention from both theoretical and industrial application point of view. Different from their molecular counterparts, granular gases allow inelastic collision, in other words, break the time-reversible symmetry because of the energy dissipation. Despite this dissipative nature and its resulting nontrivial properties, the basic equation of kinetic theory, the Boltzmann equation, can be still extended to describe the granular gases with a different collision operator, namely,
\begin{equation}
\partial_t f + v \cdot \nabla_x f = Q_{\text{in}}(f,f),
\label{boltz}
\end{equation}
where $f(t,x,v)$ is the one-particle distribution function depending on the time $t \geq 0$, position $x \in \mathbb{R}^d$ and velocity $v \in \mathbb{R}^d$ with the dimension $d \geq 1$, and $Q_{\text{in}}$ is the so-called inelastic (or granular) collision operator, whose exact expression is presented in later discussions. For convenience, we use $Q$ instead of $Q_{\text{in}}$ in the rest of this paper, whose notation should not be confused with the usual elastic collision operator. Another widely used model, first introduced by van Noije and Ernst \cite{noije}, is the spatial homogeneous inelastic Boltzmann equation based on Enskog-Boltzmann model with a heat source:
\begin{equation} \label{inBoltz}
\partial_tf-\varepsilon \Delta_vf=Q(f,f),
\end{equation}
where the distribution function $f$ depends only on the time $t$ and velocity $v$, and the term $\varepsilon \Delta_vf$ represents the diffusion effects with the diffusion coefficient $\epsilon \ll 1$, incurred by a heat bath of infinite temperature.

The numerical difficulty and cost, of course, lie in the computation of the collision operator. At each time step, the construction of the inelastic collision operator requires $O(N^{2d})$ operations and $O(N^{2d})$ storage in a direct numerical scheme. As is pointed out in \cite{filbet2005}, although the loss term in the inelastic collision can be evaluated only in $O(N^d \log N)$ operations thanks to its convolution structure, the cost of the gain part is still rather formidable. A natural question of course remains to reduce the computational cost for the entire collision operator, as well as the numerical storage -- in order words, to fully exploit the structure of the granular collision operator. In this paper, we propose a fast spectral algorithm for the inelastic collision operator, inspired by a sequence of studies on elastic Boltzmann operator \cite{fast1999, mouhot2006fast, hu2012fast, gamba_hu}. To be specific, in contrast to a direct spectral solver, in 3D this algorithm reduces the computational cost from $O(N^{6})$ to $O\left(MN^{4}\log(N)\right)$ and the storage from $O(N^{6})$ to $O(MN^{4})$, where $M \ll N^{2}$ is the number of quadrature points on $\mathbb{S}^{2}$.

@@

\section{Inelastic collision operator and its Enskog-Boltzmann equation}
\subsection{Inelastic collision }
To describe the inelastic binary collision, a reconstitution coefficient $e$ is introduced. Specifically, assuming two particles with velocities $v$ and $v_*$ are going to collide, after the collision, the velocities denoted by $v'$ and $v_*'$ are given by the so-called $\omega$-representation \cite{Villani2006} 
\begin{align}\label{omega}
\left\{
\begin{array}{l}
\displaystyle v'=v-\frac{1+e}{2}[(v-v_*)\cdot \omega ]\omega, \\[8pt]
\displaystyle v_*'=v_*+\frac{1+e}{2}[(v-v_*)\cdot \omega]\omega,
\end{array}\right.
\end{align}
where $\omega \in S^{d-1}$ is the impact direction, and $0 \leq e \leq 1$ is the restitution coefficient (with $e=1$ recovering the elastic case). It follows that
\begin{equation}
(v'-v_*')\cdot \omega=-e [(v-v_*)\cdot \omega].
\end{equation}
Furthermore, instead of the conservation of both momentum and energy in the usual molecular gases, here one has the conservation of momentum and the loss of energy:
\begin{equation} \label{lossenergy}
v'+v_*'=v+v_*; \quad v'^2+v_*'^2=v^2+v_*^2-\frac{1-e^2}{2}[(v-v_*)\cdot \omega]^2.
\end{equation}
Now we are ready to define the inelastic collision operator in its weak form
\begin{equation} \label{col_omega}
\int Q(f,f)(v)\phi(v)\diff{v}= \frac{1}{2} \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{\mathbb{S}^{d-1}} B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\left(\phi'+\phi_*'-\phi-\phi_* \right)\diff{\omega} \diff{v} \diff{v_*},
\end{equation}
where $g := v - v_\ast$, $\hat{g}$ denotes the unit vector in the direction of $g$, $\phi(t,v)$ a regular test function, $B_{\omega}$ the binary collision kernel in the $\omega$-representation, and we use the shorthands $f$ for $f(t,v)$, $f_\ast$ for $f_\ast(t, v_\ast)$ and so on. This operator clearly has collisional invariants $1$ and $v_i$ ($i = 1, 2, 3$), namely,
\begin{equation}
  \int_{\mathbb{R}^3} Q(f, f)(v) \diff v = \int_{\mathbb{R}^3} Q(f,f)(v)v_i\diff v = 0,
\end{equation}
where $v_i$ denotes the $i$-th component of the velocity $v$.

@@

Consider the inelastic Boltzmann equation with a heating source \eqref{inBoltz}. We shall revisit a few properties of this equation. First, for a solution $f$ to \eqref{inBoltz}, the conservation of mass and momentum becomes straightfoward taking $\phi = 1$ and $\phi = v$ respectively in \eqref{col_sigma}. In other words,
\begin{equation}
\rho := \int f\diff{v}\equiv \rho_0, \quad  u :=\frac{1}{\rho}\int fv \diff{v}\equiv u_0.
\end{equation}
where $\rho$ is the density and $u$ the bulk velocity. Similarly, define the kinetic energy as
\begin{equation}
  E := \frac{1}{2}\int f |v|^2\diff v
\end{equation}
and take $\phi = |v|^2$, one obtains the relation of energy dissipation
\begin{equation}\label{ki_energy}
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}\iiint B_{\sigma}(|g|,\sigma\cdot \hat{g})|g|^2(1-\sigma\cdot \hat{g})ff_*\diff{\sigma}\diff{v}\diff{v_*},
\end{equation}
where we used (\ref{lossenergy}) and (\ref{relation}). Consider the following collision kernel, namely, VHS (variable hard sphere) model
\begin{equation} \label{kernel}
B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\lambda}|g|^{\lambda}b_{\lambda}(\sigma\cdot \hat{g}),
\end{equation}
where $C_{\lambda}$ is some constant and $b_{\lambda}$ is some function. For Maxwell molecule, i.e., $\lambda=0$ in (\ref{kernel}), equation \eqref{ki_energy} becomes
\begin{equation} 
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}C_0\iiint |g|^{2}b_0(\sigma\cdot \hat{g})(1-\sigma\cdot \hat{g})ff_*\,\diff{\sigma}\,\diff{v}\,\diff{v_*}.
\end{equation}
Note that
\begin{equation}
\int C_0b_0(\sigma\cdot \hat{g})(1-\sigma\cdot \hat{g})\,\diff{\sigma}
\end{equation}
is a constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be $1$. Then
\begin{align} 
  \partial_t E -\varepsilon d\rho_0 &=-\frac{1-e^2}{16}\iint |g|^{2}ff_*\,\diff{v}\,\diff{v_*} \notag\\
   & =-\frac{1-e^2}{16}\iint (v^2+v_*^2-2v\cdot v_*)ff_*\,\diff{v}\,\diff{v_*} \notag \\
   &=-\frac{1-e^2}{8}\left(2\rho_0E - (\rho_0 u_0)^2\right).
\end{align}

@@

In 2D VHS  case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
\begin{equation}
  \int_{S^1}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\diff{\sigma}=2\pi C_{\gamma}|g|^{\gamma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot g}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}|l+m||g|\right)-1\right],
\end{equation}
then,
\begin{equation}
  G(l,m)=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho (l+m)\cdot \sigma}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}|l+m|\rho\right)-1\right],
\end{equation}
where $w_\rho$ and $w_\sigma$ are the weights of quadratures for $\rho$ and $\sigma$ respectively. $J_0$ is $0$-th order Bessel function. Therefore,
\begin{equation} 
  \hat{Q}_k=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho k\cdot \sigma}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}\rho |k|\right)-1\right]\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right],
\end{equation} 
or the loss term can be computed separately as
\begin{equation} 
\hat{Q}_k^-=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right]=\sum_{\rho}w_{\rho}4\pi^2 C_{\gamma}\rho^{\gamma+1}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[\text{J}_0\left(\frac{\pi}{L}\rho |m|\right)\hat{f}_m\right].
\end{equation}
We refer the later method as ``separate'' method and previous one as ``full'' method correspondingly.

@@

In 3D VHS case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
\begin{equation}
\int_{S^2}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\,\diff{\sigma}=4\pi C_{\gamma}|g|^{\gamma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot g}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}|l+m||g|\right)-1\right],
\end{equation}
then,
\begin{equation}
G(l,m)=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho (l+m)\cdot \sigma}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}|l+m|\rho\right)-1\right],
\end{equation}
therefore,
\begin{equation} 
\hat{Q}_k=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho k\cdot \sigma}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}\rho |k|\right)-1\right]\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right],
\end{equation} 
or the loss term can be computed separately as
\begin{equation} 
\hat{Q}_k^-=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right]=\sum_{\rho}w_{\rho}16\pi^2 C_{\gamma}\rho^{\gamma+2}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[\text{Sinc}\left(\frac{\pi}{L}\rho |m|\right)\hat{f}_m\right],
\end{equation}
which are the ``full" and ``separate" methods in 3D case.

\begin{rmk}
	In our numerical implementation, we use the Gauss-Legendre quadrature in the radial direction $\rho$, while for the integration in $\sigma$ we propose to use the spherical design in 3D case which is the near optimal quadrature on the sphere and requires fewer quadrature points than tensor product based Gauss quadratures for a large class of functions (2D case is easy to treat by using traditional quadratures). To gain a better idea of how this quadrature works, a error test of this quadrature will be given in next section.
\end{rmk}

@@

In the following numerical simulations, we use our fast spectral solver to evaluate the collision operator, paired with a standard fourier spectral method to treat the diffusion term $\varepsilon\Delta f$ and the SSP-RK3 (the 3rd order Strong Stability-Preserving Runge-Kutta method) \cite{shuTadmor} for time discretizations.

\subsection{Maxwell molecules - 2D example 1}

@@

% \begin{figure}[htb]
%   \centering
%   \includegraphics[width = .48\linewidth]{figs/sep/dt_2d_bkw_e=02}\hfill
%   \includegraphics[width = .48\linewidth]{figs/full/dt_2d_bkw_e_full=02}
%   \caption{Convergence of RK3 with $e = 0.2$. Left: separate. Right: full.}
%   \label{dt_bkw_e02}
% \end{figure}
% \begin{figure}[htb]
%   \centering
%   \includegraphics[width=0.5\textwidth]{figs/sep/dt_2d_bkw_e=05}\hfill
%   \includegraphics[width=0.5\textwidth]{figs/full/dt_2d_bkw_e_full=05}
%   \caption{Convergence of RK3 with $e = 0.5$. Left: separate. Right: full.}
%   \label{dt_bkw_e05}
% \end{figure}
% \begin{figure}[htb]
%   \centering
%   \includegraphics[width=0.5\textwidth]{figs/sep/dt_2d_bkw_e=08}\hfill
%   \includegraphics[width=0.5\textwidth]{figs/full/dt_2d_bkw_e_full=08}
%   \caption{Convergence of RK3 with $e = 0.8$. Left: separate. Right: full.}
%   \label{dt_bkw_e08}
% \end{figure}

\begin{figure}[H]
  \centering
  \subfloat[$e = 0.2$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=02}}\hfill
  \subfloat[$e = 0.2$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=02}} \\
  \subfloat[$e = 0.5$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=05}}\hfill
  \subfloat[$e = 0.5$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=05}} \\
  \subfloat[$e = 0.8$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=08}}\hfill
  \subfloat[$e = 0.8$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=08}}
  \caption{Convergence test in time of the 2D example 1 with $N = 32$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
  \label{dt_conv_1}
\end{figure}

>>>>

% personal packages
\usepackage{amsmath,amssymb}
\usepackage{subfig}
\usepackage{tabu}
\usepackage{booktabs}

@@

\title{A fast spectral method for inelastic collision operator and the heated granular flow\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
\author{Elsevier\fnref{myfootnote}}
\address{Radarweg 29, Amsterdam}
\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
\ead[url]{www.elsevier.com}

\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{support@elsevier.com}

\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
\address[mysecondaryaddress]{360 Park Avenue South, New York}

@@

\begin{keyword}
\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
\MSC[2010] 00-01\sep  99-00
\end{keyword}

@@

It has been found in the past few decays that the granular gases behave fundamentally different from the usual molecular gases modelled as elastically colliding spheres. The rich phenomenology of such systems, such as the formation of clusters and shear instability, draws a lot of attention from both theoretical and industrial application point of view. Different from their molecular counterparts, granular gases allow inelastic collision, in other words, break the time-reversible symmetry because of the energy dissipation. Despite this dissipative nature and its resulting nontrivial properties, the basic equation of kinetic theory, the Boltzmann equation, can be still extended to describe the granular gases with a different collision operator, namely,
\begin{equation}
\partial_t f + v \cdot \nabla_x f = Q_{\text{in}}(f,f),
\label{boltz}
\end{equation}
where $f(t,x,v)$ is the one-particle distribution function depending on the time $t \geq 0$, position $x \in \mathbb{R}^d$ and velocity $v \in \mathbb{R}^d$ with the dimension $d \geq 1$, and $Q_{\text{in}}$ is the so-called inelastic (or granular) collision operator, whose exact expression is presented in later discussions. For convenience, we use $Q$ instead of $Q_{\text{in}}$ in the rest of this paper, whose notation should not be confused with the usual elastic collision operator. Another widely used model, first introduced by van Noije and Ernst \cite{}, is the spatial homogeneous inelastic Boltzmann equation based on Enskog-Boltzmann model with a heat source:
\begin{equation} \label{inBoltz}
\partial_tf-\varepsilon \Delta_vf=Q(f,f),
\end{equation}
where the distribution function $f$ depends only on the time $t$ and velocity $v$, and the term $\varepsilon \Delta_vf$ represents the diffusion effects with the diffusion coefficient $\epsilon \ll 1$, incurred by a heat bath of infinite temperature.

The numerical difficulty and cost, of course, lie in the computation of the collision operator. At each time step, the construction of the inelastic collision operator requires $O(N^{2d})$ operations and $O(N^{2d})$ storage in a direct numerical scheme. As is pointed out in \cite{filbet}, although the loss term in the inelastic collision can be evaluated only in $O(N^d \log N)$ operations thanks to its convolution structure, the cost of the gain part is still rather formidable. A natural question of course remains to reduce the computational cost for the entire collision operator, as well as the numerical storage -- in order words, to fully exploit the structure of the granular collision operator. In this paper, we propose a fast spectral algorithm for the inelastic collision operator, inspired by a sequence of studies on elastic Boltzmann operator \cite{gambaHu, lexingHu, , }. To be specific, in contrast to a direct spectral solver, in 3D this algorithm reduces the computational cost from $O(N^{6})$ to $O\left(MN^{4}\log(N)\right)$ and the storage from $O(N^{6}$ to $O(MN^{4})$, where $M \ll N^{2}$ is the number of quadrature points on $\mathbb{S}^{2}$.

@@

\section{Inelastic collision operator and its Enskog-Boltzmann equation}
\subsection{Inelastic collision }
To describe the inelastic binary collision, a reconstitution coefficient $e$ is introduced. Specifically, assuming two particles with velocities $v$ and $v_*$ are going to collide, after the collision, the velocities denoted by $v'$ and $v_*'$ are given by the so-called $\omega$-representation \cite{vilani} 
\begin{align}\label{omega}
\left\{
\begin{array}{l}
\displaystyle v'=v-\frac{1+e}{2}[(v-v_*)\cdot \omega ]\omega, \\[8pt]
\displaystyle v_*'=v_*+\frac{1+e}{2}[(v-v_*)\cdot \omega]\omega,
\end{array}\right.
\end{align}
where $\omega \in S^{d-1}$ is the impact direction, and $0 \leq e \leq 1$ is the restitution coefficient (with $e=1$ recovering the elastic case). It follows that
\begin{equation}
(v'-v_*')\cdot \omega=-e [(v-v_*)\cdot \omega].
\end{equation}
Furthermore, instead of the conservation of both momentum and energy in the usual molecular gases, here one has the conservation of momentum and the loss of energy:
\begin{equation} \label{lossenergy}
v'+v_*'=v+v_*; \quad v'^2+v_*'^2=v^2+v_*^2-\frac{1-e^2}{2}[(v-v_*)\cdot \omega]^2.
\end{equation}
Now we are ready to define the inelastic collision operator in its weak form
\begin{equation} \label{col_omega}
\int Q(f,f)(v)\phi(v)\diff{v}= \frac{1}{2} \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{\mathbb{S}^{d-1}} B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\left(\phi'+\phi_*'-\phi-\phi_* \right)\diff{\omega} \diff{v} \diff{v_*},
\end{equation}
where $g := v - v_\ast$, $\hat{g}$ denotes the unit vector in the direction of $g$, $\phi(t,v)$ a regular test function, $B_{\omega}$ the binary collision kernel in the $\omega$-representation, and we use the shorthands $f$ for $f(t,v)$, $f_\ast$ for $f_\ast(t, v_\ast)$ and so on. This operator clearly has collisional invariants $1$ and $v_i$ ($i = 1, 2, 3$), namely,
\begin{equation}
  \int_{\mathbb{R}^3} Q(f, f)(v) \diff v = \int_{\mathbb{R}^3} Q(f,f)(v)v_i\diff v = 0,
\end{equation}
where $v_i$ denotes the $i$-th component of the velocity $v$.

@@

Consider the inelastic Boltzmann equation with a heating source \eqref{inBoltz}. We shall revisit a few properties of this equation. First, for a solution $f$ to \eqref{inBoltz}, the conservation of mass and momentum becomes straightfoward taking $\phi = 1$ and $\phi = v$ respectively in \eqref{col_sigma}. In other words,
\begin{equation}
\rho := \int f\diff{v}\equiv \rho_0, \quad  u :=\frac{1}{\rho}\int fv \diff{v}\equiv u_0.
\end{equation}
where $\rho$ is the density and $u$ the bulk velocity. Similarly, define the kinetic energy as
\begin{equation}
  E := \frac{1}{2}\int f |v|^2\diff v
\end{equation}
and take $\phi = |v|^2$, one obtains the relation of energy dissipation
\begin{equation}\label{ki_energy}
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}\iiint B_{\sigma}(|g|,\sigma\cdot \hat{g})|g|^2(1-\sigma\cdot \hat{g})ff_*\diff{\sigma}\diff{v}\diff{v_*},
\end{equation}
where we used (\ref{lossenergy}) and (\ref{relation}). Consider the following collision kernel
\begin{equation} \label{kernel}
B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\lambda}|g|^{\lambda}b_{\lambda}(\sigma\cdot \hat{g}),
\end{equation}
where $C_{\lambda}$ is some constant and $b_{\lambda}$ is some function. For Maxwell molecule, i.e., $\lambda=0$ in (\ref{kernel}), equation \eqref{ki_energy} becomes
\begin{equation} 
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}C_0\iiint |g|^{2}b_0(\sigma\cdot \hat{g})(1-\sigma\cdot \hat{g})ff_*\,\diff{\sigma}\,\diff{v}\,\diff{v_*}.
\end{equation}
Note that
\begin{equation}
\int C_0b_0(\sigma\cdot \hat{g})(1-\sigma\cdot \hat{g})\,\diff{\sigma}
\end{equation}
is a constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be $1$. Then
\begin{align} 
  \partial_t E -\varepsilon d\rho_0 &=-\frac{1-e^2}{16}\iint |g|^{2}ff_*\,\diff{v}\,\diff{v_*} \notag\\
   & =-\frac{1-e^2}{16}\iint (v^2+v_*^2-2v\cdot v_*)ff_*\,\diff{v}\,\diff{v_*} \notag \\
   &=-\frac{1-e^2}{8}\left(2\rho_0E - (\rho_0 u_0)^2\right).
\end{align}

@@

In 2D, and VHS case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
\begin{equation}
  \int_{S^1}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\,\diff{\sigma}=2\pi C_{\gamma}|g|^{\gamma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot g}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}|l+m||g|\right)-1\right],
\end{equation}
then let $\rho=|g|$, $\sigma=\hat{g}$,
\begin{equation}
  G(l,m)=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho (l+m)\cdot \sigma}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}|l+m|\rho\right)-1\right],
\end{equation}
therefore,
\begin{equation} 
  \hat{Q}_k=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho k\cdot \sigma}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}\rho |k|\right)-1\right]\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right],
\end{equation} 
or the loss term can be computed separately as
\begin{equation} 
\hat{Q}_k^-=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right]=\sum_{\rho}w_{\rho}4\pi^2 C_{\gamma}\rho^{\gamma+1}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[\text{J}_0\left(\frac{\pi}{L}\rho |m|\right)\hat{f}_m\right].
\end{equation}
We refer the later method as "seperate" method and previous one as "full" method correspondingly.

@@

In 3D, and VHS case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
\begin{equation}
\int_{S^2}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\,\diff{\sigma}=4\pi C_{\gamma}|g|^{\gamma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot g}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}|l+m||g|\right)-1\right],
\end{equation}
then let $\rho=|g|$, $\sigma=\hat{g}$,
\begin{equation}
G(l,m)=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho (l+m)\cdot \sigma}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}|l+m|\rho\right)-1\right],
\end{equation}
therefore,
\begin{equation} 
\hat{Q}_k=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho k\cdot \sigma}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}\rho |k|\right)-1\right]\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right],
\end{equation} 
or the loss term can be computed separately as
\begin{equation} 
\hat{Q}_k^-=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right]=\sum_{\rho}w_{\rho}16\pi^2 C_{\gamma}\rho^{\gamma+2}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[\text{Sinc}\left(\frac{\pi}{L}\rho |m|\right)\hat{f}_m\right],
\end{equation}
which are the "full" and "separate" methods in 3D case.

@@

In the following numerical simulations, we use our fast spectral solver to evaluate the collision operator, paired with a standard fourier spectral method to treat the diffusion term $\varepsilon\Delta f$ and the SSP-RK3 (the 3rd order Strong Stability-Preserving Runge-Kutta method) \cite{ShuTadmor} for time discretizations.
\subsection{2D examples - Maxwell molecules}

@@

In Figure~\ref{dt_conv_1} we plot the relation between the errors versus different $\Delta t$s for $e = 0.2, 0.5$ and $0.8$, $T_\text{final}=2$. A third order convergence rate can be seen very easily. Since in this test we fix all the parameters in our spectral method as $\Delta t$ changes, these figures also imply that the error of the spectral method computing the collision is really small, or at least around $O(10^{-9})$.
% \begin{figure}[htb]
%   \centering
%   \includegraphics[width = .48\linewidth]{figs/sep/dt_2d_bkw_e=02}\hfill
%   \includegraphics[width = .48\linewidth]{figs/full/dt_2d_bkw_e_full=02}
%   \caption{Convergence of RK3 with $e = 0.2$. Left: separate. Right: full.}
%   \label{dt_bkw_e02}
% \end{figure}
% \begin{figure}[htb]
%   \centering
%   \includegraphics[width=0.5\textwidth]{figs/sep/dt_2d_bkw_e=05}\hfill
%   \includegraphics[width=0.5\textwidth]{figs/full/dt_2d_bkw_e_full=05}
%   \caption{Convergence of RK3 with $e = 0.5$. Left: separate. Right: full.}
%   \label{dt_bkw_e05}
% \end{figure}
% \begin{figure}[htb]
%   \centering
%   \includegraphics[width=0.5\textwidth]{figs/sep/dt_2d_bkw_e=08}\hfill
%   \includegraphics[width=0.5\textwidth]{figs/full/dt_2d_bkw_e_full=08}
%   \caption{Convergence of RK3 with $e = 0.8$. Left: separate. Right: full.}
%   \label{dt_bkw_e08}
% \end{figure}

\begin{figure}[htb]
  \centering
  \subfloat[$e = 0.2$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=02}}\hfill
  \subfloat[$e = 0.2$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=02}} \\
  \subfloat[$e = 0.5$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=05}}\hfill
  \subfloat[$e = 0.5$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=05}} \\
  \subfloat[$e = 0.8$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=08}}\hfill
  \subfloat[$e = 0.8$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=08}}
  \caption{convergence in time 1}
  \label{dt_conv_1}
\end{figure}
